---
title: "북한의 가뭄 변동과 지역별 쌀 가격의 관계: 평양·신의주·혜산 월별 패널자료를 이용한 실증분석"
author: "정주희"
date: today
lang: ko

format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: false
    code-fold: true
    code-summary: "코드 보기"
    theme: paper

execute:
  echo: true
  warning: false
  message: false
---

# 1. Initial setting {#top}

## Loading packages

```{r}
#| label: setup

source("code/00_initial.R")
```

[Back to top](#top)


# 2. Data preparation

## 2.1 Data import and preprocessing

```{r}
#| label: data-prep                 
#| include: false                   

source("code/01_01_data_weather_rice.R")  
source("code/01_02_asiapress.R")       
```


### Precipitation data

The daily precipitation data cover 27 weather stations from January 1973 to July 2026. The precipitation variable has a missing rate of 66.28%. Missing precipitation values are addressed in scripts `02_00`–`02_07` before aggregation to the monthly level.

```{r}
#| echo: true

precip_summary <- tibble(
  item = c(
    "Number of weather stations",
    "Start date",
    "End date",
    "Total observations",
    "Observed precipitation",
    "Missing precipitation",
    "Missing rate (%)",
    "Zero precipitation",
    "Positive precipitation"
  ),
  value = c(
    n_distinct(weather_daily_raw$station),
    as.character(min(weather_daily_raw$date, na.rm = TRUE)),
    as.character(max(weather_daily_raw$date, na.rm = TRUE)),
    format(
      nrow(weather_daily_raw),
      big.mark = ",",
      scientific = FALSE
    ),
    format(
      sum(!is.na(weather_daily_raw$precipitation_total_day)),
      big.mark = ",",
      scientific = FALSE
    ),
    format(
      sum(is.na(weather_daily_raw$precipitation_total_day)),
      big.mark = ",",
      scientific = FALSE
    ),
    round(
      mean(is.na(weather_daily_raw$precipitation_total_day)) * 100,
      2
    ),
    format(
      sum(
        weather_daily_raw$precipitation_total_day == 0,
        na.rm = TRUE
      ),
      big.mark = ",",
      scientific = FALSE
    ),
    format(
      sum(
        weather_daily_raw$precipitation_total_day > 0,
        na.rm = TRUE
      ),
      big.mark = ",",
      scientific = FALSE
    )
  )
)

knitr::kable(
  precip_summary,
  col.names = c("Item", "Value"),
  align = c("l", "c"),
  caption = "Summary of daily precipitation data"
)
```


### Rice price data (Daily NK)

Rice price observations for Pyongyang, Sinuiju, and Hyesan are aggregated to monthly averages. The data cover January 2013 to July 2026. Months without observations are retained as missing.


```{r}
#| echo: true

rice_summary <- tibble(
  item = c(
    "Number of original observations",
    "Start date",
    "End date",
    "Number of total months",
    "Number of observed months",
    "Number of missing months"
  ),
  value = c(
    nrow(rice_price_obs),
    as.character(min(rice_price_obs$date, na.rm = TRUE)),
    as.character(max(rice_price_obs$date, na.rm = TRUE)),
    nrow(all_months),
    nrow(rice_price_monthly),
    nrow(missing_rice_months)
  )
)

knitr::kable(
  rice_summary,
  col.names = c("Item", "Value"),
  align = c("l", "c"),
  caption = "Summary of rice price data"
)
```


Months without rice price observations are shown below. These months are retained as missing and are not interpolated.

```{r}
#| echo: true

missing_rice_months |>
  knitr::kable(
    col.names = "Missing month",
    align = "c",
    caption = "Months without rice price observations"
  )
```


The following table reports descriptive statistics for monthly mean rice prices in the three study locations.

```{r}
#| echo: true

rice_descriptive <- rice_price_monthly |>
  select(
    rice_pyongyang,
    rice_sinuiju,
    rice_hyesan
  ) |>
  pivot_longer(
    everything(),
    names_to = "region",
    values_to = "rice_price"
  ) |>
  mutate(
    region = recode(
      region,
      rice_pyongyang = "Pyongyang",
      rice_sinuiju   = "Sinuiju",
      rice_hyesan    = "Hyesan"
    )
  ) |>
  group_by(region) |>
  summarise(
    N    = sum(!is.na(rice_price)),
    Mean = mean(rice_price, na.rm = TRUE),
    SD   = sd(rice_price, na.rm = TRUE),
    Min  = min(rice_price, na.rm = TRUE),
    Max  = max(rice_price, na.rm = TRUE),
    .groups = "drop"
  )

knitr::kable(
  rice_descriptive,
  digits = 1,
  col.names = c(
    "Region",
    "N",
    "Mean",
    "SD",
    "Min",
    "Max"
  ),
  align = c("l", "c", "c", "c", "c", "c"),
  caption = "Descriptive statistics of monthly rice prices"
)
```


The monthly series shows the evolution of rice prices across Pyongyang, Sinuiju, and Hyesan, with a marked acceleration in prices beginning in June 2025 and a particularly sharp increase in August 2025.

```{r}
#| echo: true

rice_price_monthly_plot
```

For descriptive purposes, months in which rice prices increased by 20% or more from the preceding month are summarized below.

```{r}
#| echo: true

knitr::kable(
  rice_price_increase_table,
  col.names = c(
    "Month",
    "Pyongyang",
    "Sinuiju",
    "Hyesan"
  ),
  align = c("l", "c", "c", "c"),
  caption = "Monthly rice price increases of 20% or more"
)
```

[Back to top](#top)


## Supplementary market data: AsiaPress

AsiaPress provides observations on rice, corn, fuel prices, and exchange rates in North Korean markets. 
Because the survey locations vary over time, these data do not represent a fixed regional price series or a national average. 
They are therefore used as supplementary evidence on broader market price movements rather than as the main dependent variable.

The monthly analysis covers April 2017 through August 2026. Months with no survey observations are retained explicitly as missing.

```{r}
#| label: tbl-asiapress-summary
#| echo: true

asiapress_summary <- tibble(
  item = c(
    "Collection date",
    "Observation start date",
    "Observation end date",
    "Monthly analysis start",
    "Monthly analysis end",
    "Cleaned observations",
    "Number of monthly observations",
    "Months with no survey observations"
  ),

  value = c(
    as.character(collection_date),

    as.character(
      min(asiapress_data_clean$date, na.rm = TRUE)
    ),

    as.character(
      max(asiapress_data_clean$date, na.rm = TRUE)
    ),

    min(asiapress_monthly$month),

    max(asiapress_monthly$month),

    nrow(asiapress_data_clean),

    nrow(asiapress_monthly),

    sum(asiapress_monthly$n_obs == 0)
  )
)

knitr::kable(
  asiapress_summary,
  col.names = c("Item", "Value"),
  align = c("l", "c"),
  caption = "Summary of AsiaPress market data"
)
```


The following figure compares the monthly movements of rice and corn prices reported by AsiaPress.

```{r}
#| label: fig-asiapress-rice-corn
#| echo: true
#| fig-width: 9
#| fig-height: 5.5
#| fig-cap: "Monthly rice and corn prices reported by AsiaPress"

asiapress_rice_corn_plot
```


To examine the recent food-price surge, the largest year-on-year increases in rice and corn prices are reported below.

```{r}
#| label: tbl-asiapress-price-surge
#| echo: true

knitr::kable(
  recent_price_surge_table_display,

  col.names = c(
    "Commodity",
    "Month",
    "Current price",
    "Price 12 months earlier",
    "YoY change (%)"
  ),

  align = c(
    "l",
    "c",
    "c",
    "c",
    "c"
  ),

  caption = paste0(
    "Largest year-on-year increases in rice and corn prices, ",
    recent_period_label
  )
)
```


The AsiaPress series shows that rice prices entered a pronounced upward phase before the sharp increase in corn prices observed in 2026. 
This temporal pattern is used only as descriptive evidence and does not by itself imply a causal relationship between rice and corn prices.


[Back to top](#top)

## 2.2 Missing precipitation treatment


### 2.2.1 Precipitation without imputation

```{r}
#| label: missing-precip-raw

source("code/02_01_precip_raw.R")
```

[Back to top](#top)


### 2.2.2 Temperature-based zero replacement

```{r}
#| label: missing-precip-zero

source("code/02_02_precip_zero.R")
```

[Back to top](#top)



### 2.2.3 Precipitation with multiple imputation 

```{r}
#| label: missing-precip-mi

source("code/02_03_precip_mi.R")
```

[Back to top](#top)


### 2.2.4 Precipitation with Random Forest

```{r}
#| label: missing-precip-rf

source("code/02_04_precip_rf.R")
```

[Back to top](#top)


### 2.2.5 Precipitation with XGBoost
```{r}
#| label: missing-precip-xgb

source("code/02_05_precip_xgb.R")
```

[Back to top](#top)


### 2.2.6 Validation and comparison of Precipitation treatments 

```{r}
#| label: missing-precip-val

source("code/02_06_precip_validation.R")
```

[Back to top](#top)


### 2.2.7 Missing Precipitation Treatment Summary

```{r}
#| label: missing-precip

source("code/02_07_missing_precip.R")
```

[Back to top](#top)


## 2.3 Drought index construction

### 2.3.1 SPI

```{r}
#| label: spi

source("code/03_01_spi.R")
```

[Back to top](#top)


### 2.3.2 SPEI

```{r}
#| label: spei

source("code/03_02_spei.R")
```

[Back to top](#top)


### 2.3.3 SMI

```{r}
#| label: smi

source("code/03_03_smi.R")
```

[Back to top](#top)


### 2.3.4 Drought Indices Summary

```{r}
#| label: drought-index

source("code/03_04_drought_indices_summary.R")
```

[Back to top](#top)


## 2.4 Final dataset

```{r}
#| label: final-dataset

source("code/04_final_dataset.R")
```

[Back to top](#top)


# 3. Data review

## 3.1 Descriptive statistics

```{r}
#| label: descriptive

source("code/05_00_descriptive.R")
```

[Back to top](#top)


# 4. Estimation


## 4.1 Pooled estimation

```{r}
#| label: pooled-estimation

source("code/06_01_pooled_estimation.R")
```

[Back to top](#top)


## 4.2 Region-specific estimation

```{r}
#| label: region-estimation

source("code/06_02_region_estimation.R")
```

[Back to top](#top)



## 4.3 Lagged effects

```{r}
#| label: lagged-effects

source("code/06_03_lagged_effects.R")
```

[Back to top](#top)



## 4.4 Robustness checks

```{r}
#| label: robustness

source("code/06_04_robustness.R")
```

[Back to top](#top)


## 4.5 Estimation summary 

```{r}
#| label: estimation

source("code/06_05_estimation_summary.R")
```

[Back to top](#top)


# 5. Visualization

## 5.1 Figure 1. xxx

```{r}
#| label: fig1
source("code/07_00_plot.R")
```

## 5.2 Figure 2. xxx

```{r}
#| label: fig2

```

[Back to top](#top)
