---
title: "북한의 가뭄 변동과 지역별 쌀 가격의 관계: 평양·신의주·혜산 월별 패널자료를 이용한 실증분석"
author: "정주희"
date: today
lang: ko

format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: false
    code-fold: true
    code-summary: "코드 보기"
    theme: paper

execute:
  echo: true
  warning: false
  message: false
---

# 1. Initial setting {#top}

## Loading packages

```{r}
#| label: setup

source("code/00_initial.R")
```

[Back to top](#top)


# 2. Data preparation

## 2.1 Data import and preprocessing

```{r}
#| label: data-prep

source("code/01_data.R")
```

[Back to top](#top)


## 2.2 Missing precipitation treatment


### 2.2.1 Precipitation without imputation

```{r}
#| label: missing-precip-raw

source("code/02_01_precip_raw.R")
```

[Back to top](#top)


### 2.2.2 Temperature-based zero replacement

```{r}
#| label: missing-precip-zero

source("code/02_02_precip_zero.R")
```

[Back to top](#top)



### 2.2.3 Precipitation with multiple imputation 

```{r}
#| label: missing-precip-mi

source("code/02_03_precip_mi.R")
```

[Back to top](#top)


### 2.2.4 Precipitation with Random Forest

```{r}
#| label: missing-precip-rf

source("code/02_04_precip_rf.R")
```

[Back to top](#top)


### 2.2.5 Precipitation with XGBoost
```{r}
#| label: missing-precip-xgb

source("code/02_05_precip_xgb.R")
```

[Back to top](#top)


### 2.2.6 Validation and comparison of Precipitation treatments 

```{r}
#| label: missing-precip-val

source("code/02_06_precip_validation.R")
```

[Back to top](#top)


### 2.2.7 Missing Precipitation Treatment Summary

```{r}
#| label: missing-precip

source("code/02_07_missing_precip.R")
```

[Back to top](#top)


## 2.3 Drought index construction

### 2.3.1 SPI

```{r}
#| label: spi

source("code/03_01_spi.R")
```

[Back to top](#top)


### 2.3.2 SPEI

```{r}
#| label: spei

source("code/03_02_spei.R")
```

[Back to top](#top)


### 2.3.3 SMI

```{r}
#| label: smi

source("code/03_03_smi.R")
```

[Back to top](#top)


### 2.3.4 Drought Indices Summary

```{r}
#| label: drought-index

source("code/03_04_drought_indices_summary.R")
```

[Back to top](#top)


## 2.4 Final dataset

```{r}
#| label: final-dataset

source("code/04_final_dataset.R")
```

[Back to top](#top)


# 3. Data review

## 3.1 Descriptive statistics

```{r}
#| label: descriptive

source("code/05_00_descriptive.R")
```

[Back to top](#top)


# 4. Estimation


## 4.1 Pooled estimation

```{r}
#| label: pooled-estimation

source("code/06_01_pooled_estimation.R")
```

[Back to top](#top)


## 4.2 Region-specific estimation

```{r}
#| label: region-estimation

source("code/06_02_region_estimation.R")
```

[Back to top](#top)



## 4.3 Lagged effects

```{r}
#| label: lagged-effects

source("code/06_03_lagged_effects.R")
```

[Back to top](#top)



## 4.4 Robustness checks

```{r}
#| label: robustness

source("code/06_04_robustness.R")
```

[Back to top](#top)


## 4.5 Estimation summary 

```{r}
#| label: estimation

source("code/06_05_estimation_summary.R")
```

[Back to top](#top)


# 5. Visualization

## 5.1 Figure 1. xxx

```{r}
#| label: fig1
source("code/07_00_plot.R")
```

## 5.2 Figure 2. xxx

```{r}
#| label: fig2

```

[Back to top](#top)
